---
title: "Group Homework 1"
author: "Aastha, Ameer, Ananth, Karen"
output: word_document
---

# Importing required libraries for the analysis
```{r,message=FALSE,warning=FALSE}
library(dplyr)
library(stats)
library(cluster)
library(ggplot2)
library(factoextra)
```

# Reading the Wholesale customers datafile
```{r}
customers = read.csv("Wholesale customers data.csv")
```

# Boxplot for numerical columns
```{r}
boxplot(customers[3:8], col = "grey", outcol = "red")
```

# Removing outliers (1.5 * IQR)
```{r}
identify_outliers <- function(x) {
  Q1 <- quantile(x, 0.25)
  Q3 <- quantile(x, 0.75)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  return(x < lower_bound | x > upper_bound)
}

# Identify outliers for each column
outliers <- apply(customers[, 3:8], 2, identify_outliers)

# Identify rows with one or more outliers
outlier_rows <- apply(outliers, 1, any)

customers_outliers <- customers[outlier_rows, ]
customers_clean <- customers[!outlier_rows, ]
```

# Calculating distance matrix
```{r}
distance_matrix = dist(customers_clean[,3:8], method = "euclidean")
```

# Plotting SSE to observe elbow point
```{r}
SSE_curve <- c()
for (n in 1:30) {
  kcluster = kmeans(customers_clean[,3:8], n)
  sse = kcluster$tot.withinss
  SSE_curve[n] = sse}
# plot SSE against number of clusters
plot(1:30, SSE_curve, type = "b")
```

# Clustering using K-Means method
```{r}
kcluster = kmeans(customers_clean[,3:8], centers = 3)
kcluster$centers
kcluster$cluster
customers_clean$K_Cluster <- kcluster$cluster

write.csv(customers_clean,"customer_clust.csv", row.names = TRUE)
```

# Evaluating the clustering solution using Silhouette Coefficient
```{r}
sc = silhouette(customers_clean$K_Cluster, dist = distance_matrix)
summary(sc)
```

# Plotting datapoints after dimensionality reduction using PCA
```{r}
customers_clean$K_Cluster = as.factor(customers_clean$K_Cluster)
pca_result <- prcomp(customers_clean[, 3:8], scale = TRUE)
fviz_pca_ind(pca_result,
             geom.ind = "point",
             col.ind = customers_clean$K_Cluster,
             palette = "jco",
             addEllipses = TRUE,
             ellipse.level = 0.90,
             legend.title = "Cluster") +
  labs(title = "PCA of Wholesale Customers Clustering") +
  theme_minimal()
```

# Writing outliers to a CSV file
```{r}
write.csv(customers_outliers,"customer_outliers.csv", row.names = TRUE)
```